Lower Bounds on the Minimax Risk for the \\ Source Localization Problem

### 2017-05-10 paper.tex
Make minor edits; make paper size a4.

\documentclass[conference,letterpaper]{IEEEtran}

@@

For EEG source localization, we hope to find algorithms that achieve scalings
close to the bounds we present.  We also expect to build upon the techniques
presented in this paper to derive bounds for spherical and realistic brain
models.

@@

Sensors are assumed to be uniformly distributed over the domain, i.e., if there
are $m$ sensors, they are placed at locations $s = 0$, $S/m$, $2S/m$,~\dots,
$(m{-}1)S/m$ (the offset of the first sensor is arbitrary, so without loss of
generality we take it to be 0). The lower bounds we provide are for \emph{this
	specific sensor configuration}. For a discussion on why this configuration
might be an appropriate choice in the minimax setting, see
section~\ref{sec:discussion}.

@@

We consider two sensor models: a simplistic ``point sensor'' model and an
``integrator sensor'' model, to be used in sections~\ref{sec:lecam-lb}
and~\ref{sec:physical-model} respectively.

@@

	section~\ref{sec:source-localization} under the point sensor model, the

@@

tensorization of the KL-divergence~\cite{Duchi2015Information}, we see that:

@@

	\norm{P_1^n - P_0^n}_{TV}^2 \leq \frac{1}{2} D_{KL}(P_0^n \Vert P_1^n) = \frac{n}{2} D_{KL}(P_0 \Vert P_1)

@@

	D_{KL}(P_0 \Vert P_1) = (\v \mu_0 - \v \mu_1)^T \m \Sigma^{-1} (\v \mu_0 - \v \mu_1)

@@

section~\ref{sec:sensor-model}. Hence, combining

@@

	\mathfrak{M}_n(\mathcal{P}, \Phiorho) \geq \frac{\delta^2}{2} \Bigg[ 1 - \sqrt{ \frac{n}{2\sigma^2} \norm*{\v x(\theta_0) - \v x(\theta_1)}^2 } \Bigg]

@@

section~\ref{sec:signal-model}. For an impulse response $g$ which is Lipschitz

@@

once, since $w < S/2$, as stated in section~\ref{sec:signal-model}.

@@

The bound in Theorem~\ref{thm:main-lb} may mislead one to think that the error
in source localization goes to zero, as the number of sensors goes to infinity.
This observation contradicts the information-theoretic lower bound given
in~\cite{Grover2016Fundamental}, where even with an infinite number of sensors,
the lower bound is strictly greater than zero. This indicates that the bound in
section~\ref{sec:lecam-lb} is failing to capture some effect which leads to
saturation of error as the number of sensors increases.

@@

This observation motivates the ``integrator sensor'' model defined in
section~\ref{sec:sensor-model}. With this model, the bounding technique used in
section~\ref{sec:lecam-lb} no longer easily admits a closed-form solution which
gives a bound for all values of $m$. We instead compute the bound numerically,
using exact expressions for error probability in a binary hypothesis test. For
Gaussian distributions of equal variance and a uniform prior on the hypotheses,
the minimum possible error probability (over all tests) is a $Q$-function in
the distance between their means~\cite[Ch.~7]{Scharf1991Statistical},
\begin{equation}
	P_e = \inf_\psi \mathbb P(\psi(Y^n) \neq V) = Q\bigg(\frac{\norm{\v\mu_0 - \v\mu_1}}{2\sigma(m)}\bigg).
\end{equation}
where $Q(x) = \int_x^\infty \frac{1}{\sqrt{2\pi}} e^{-u^2/2} du$. Hence, from
Theorem~\ref{thm:est-to-testing},
\begin{equation}
	\mathfrak{M}_n(\mathcal{P}, \Phiorho) \geq \Phi(\delta) P_e = \delta^2 Q\bigg(\frac{\norm{\v x(\theta_0) - \v x(\theta_1)}}{2\sigma(m)}\bigg).
\end{equation}
This bound is valid for any $\{\theta_0, \theta_1\}$ constituting a
$2\delta$-packing of $\Theta$. To numerically optimize over all possible
packings, we relax the process by fixing $\theta_0 = 0$ and optimizing over
$\theta_1 = 2\delta$.  We compute sensor values for a triangular impulse
response, by performing numerical integration, and find the best possible
$\delta$ by grid search.

\begin{figure}[t]
	\centering
	\includegraphics[width=2.8in]{lb-vs-m-annotated}
	\vspace{-3pt}
	\caption{A numerically generated plot of the lower bound on minimax risk of
		source localization, against the number of uniformly distributed
		sensors, $m$, used to estimate the location (code available
		online~\cite{FullVersion}). The different curves are for different
		numbers of ``trials'', $n$ (number of i.i.d.\ sensor observations). The
		bound saturates for large $m$, indicating that for a \emph{fixed}
		number of trials, there is decreasing benefit to using more sensors.
		However, the plot suggests that with more trials, the bound saturates
		at increasingly larger numbers of sensors. Thus, there may be benefit
		to increasing the number of sensors when you have more trials.}
	\label{fig:numerical}
\end{figure}

@@

While the lower bound presented in section~\ref{sec:lecam-lb} is non-physical
for a large number of sensors, it still gives some important insights.  The
bound tells us how the error scales with the number of sensors up to $m =
\floor*{\frac{S}{c}}$, for sensors of a constant width $c$. Averaging noise
over a larger number of trials $n$ may help one achieve a target mean squared
error in location. An analytical derivation of the trade-off between $m$ and
$n$ in achieving a certain target MSE is relegated to future work. Further,
more sophisticated techniques for deriving lower bounds, such as Fano's
method~\cite[Sec.~2.7.1]{Tsybakov2009Introduction} are yet to be analyzed.

@@

While the bounds in section~\ref{sec:lecam-lb} were derived for a
one-dimensional circular domain, extensions to a 1D linear domain (of fixed
length $S$) and to a $d$-dimensional domain (circular or linear) are fairly
straightforward, under certain conditions. For a 1D linear domain, we require
that sensors capture the entirety of the impulse response even at the edges of
the domain. Hence, sensors must be placed uniformly in $[-w/2, S{+}w/2]$. For
extending to $d$ dimensions, it suffices to apply Assouad's
Lemma~\cite[Sec.~2.7.2]{Tsybakov2009Introduction} to separate the problem into
$d$ independent 1D problems. Such extensions substantially broaden the
application domains where this bound could be useful.

@@

%So far, we've relied on the presence of a circular domain for to make the
%arguments in the proof. However, the most important requirement was that
%\emph{all} parts of the impulse response should get \emph{sensed} uniformly.
%So, if $\theta$ were restricted to the region $[0, S)$ (now on a linear,
%non-periodic domain), and if sensors were available to sense the signal from
%$[-w/2, S{+}w/2)$ (for an impulse response with restricted support as described
%in section~\ref{sec:signal-model}), then the proof for sensor noise would still
%fall through, albeit with slightly different constants.

>>>>

\documentclass[conference,a4paper]{IEEEtran}

@@

For EEG source localization, we intend to find algorithms that achieve scalings
close to the bounds we present.  We also expect to build upon the techniques
presented in this paper to derive bounds for spherical and realistic brain
models.

@@

Sensors are assumed to be uniformly distributed over the domain, i.e., if there
are $m$ sensors, they are placed at locations $s = 0$, $S/m$, $2S/m$,~\dots,
$(m{-}1)S/m$ (the offset of the first sensor is arbitrary, so without loss of
generality we take it to be 0). The lower bounds we provide are for \emph{this
	specific sensor configuration}. For a discussion on why this configuration
might be an appropriate choice in the minimax setting, see
Section~\ref{sec:discussion}.

@@

We consider two sensor models: a simplistic ``point sensor'' model and an
``integrator sensor'' model, to be used in Sections~\ref{sec:lecam-lb}
and~\ref{sec:physical-model} respectively.

@@

	Section~\ref{sec:source-localization} under the point sensor model, the

@@

tensorization of the KL-divergence~\cite{Duchi2015Information}, we see that

@@

	\norm{P_1^n - P_0^n}_{TV}^2 \leq \frac{1}{2} D_{KL}(P_0^n \Vert P_1^n) = \frac{n}{2} D_{KL}(P_0 \Vert P_1).

@@

	D_{KL}(P_0 \Vert P_1) = (\v \mu_0 - \v \mu_1)^T \m \Sigma^{-1} (\v \mu_0 - \v \mu_1),

@@

Section~\ref{sec:sensor-model}. Hence, combining

@@

	\mathfrak{M}_n(\mathcal{P}, \Phiorho) \geq \frac{\delta^2}{2} \Bigg[ 1 - \sqrt{ \frac{n}{2\sigma^2} \norm*{\v x(\theta_0) - \v x(\theta_1)}^2 } \Bigg].

@@

Section~\ref{sec:signal-model}. For an impulse response $g$ which is Lipschitz

@@

once, since $w < S/2$, as stated in Section~\ref{sec:signal-model}.

@@

The bound in Theorem~\ref{thm:main-lb} may mislead one to think that the error
in source localization goes to zero, as the number of sensors goes to infinity.
This observation contradicts the information-theoretic lower bound given
in~\cite{Grover2016Fundamental}, where even with an infinite number of sensors,
the lower bound is strictly greater than zero. This indicates that the bound in
Section~\ref{sec:lecam-lb} is failing to capture some effect which leads to
saturation of error as the number of sensors increases.

@@

This observation motivates the ``integrator sensor'' model defined in
Section~\ref{sec:sensor-model}. With this model, the bounding technique used in
Section~\ref{sec:lecam-lb} no longer easily admits a closed-form solution which
gives a bound for all values of $m$. We instead compute the bound numerically,
using exact expressions for error probability in a binary hypothesis test. For
Gaussian distributions of equal variance and a uniform prior on the hypotheses,
the minimum possible error probability (over all tests) is a $Q$-function in
the distance between their means~\cite[Ch.~7]{Scharf1991Statistical},
\begin{equation}
	P_e = \inf_\psi \mathbb P(\psi(Y^n) \neq V) = Q\bigg(\frac{\norm{\v\mu_0 - \v\mu_1}}{2\sigma(m)}\bigg).
\end{equation}
where $Q(x) = \int_x^\infty \frac{1}{\sqrt{2\pi}} e^{-u^2/2} du$. Hence, from
Theorem~\ref{thm:est-to-testing},
\begin{equation}
	\mathfrak{M}_n(\mathcal{P}, \Phiorho) \geq \Phi(\delta) P_e = \delta^2 Q\bigg(\frac{\norm{\v x(\theta_0) - \v x(\theta_1)}}{2\sigma(m)}\bigg).
\end{equation}
This bound is valid for any $\{\theta_0, \theta_1\}$ constituting a
$2\delta$-packing of $\Theta$. To numerically optimize over all possible
packings, we relax the process by fixing $\theta_0 = 0$ and optimizing over
$\theta_1 = 2\delta$.  We compute sensor values for a triangular impulse
response by performing numerical integration, and find the best possible
$\delta$ by grid search.

\begin{figure}[t]
	\centering
	\includegraphics[width=2.8in]{lb-vs-m-annotated}
	\vspace{-3pt}
	\caption{A numerically generated plot of the lower bound on minimax risk of
		source localization, against the number of uniformly distributed
		sensors, $m$, following the ``integrator sensor'' model (code available
		online~\cite{FullVersion}). The different curves are for different
		numbers of ``trials'', $n$ (number of i.i.d.\ sensor observations). The
		bound saturates for large $m$, indicating that for a \emph{fixed}
		number of trials, there is decreasing benefit to using more sensors.
		However, the plot suggests that with more trials, the bound saturates
		at increasingly larger numbers of sensors. Thus, there may be benefit
		to increasing the number of sensors when you have more trials.}
	\label{fig:numerical}
\end{figure}

@@

While the lower bound presented in Section~\ref{sec:lecam-lb} is non-physical
for a large number of sensors, it still gives some important insights.  The
bound tells us how the error scales with the number of sensors up to $m =
\floor*{\frac{S}{c}}$, for sensors of a constant width $c$. Averaging noise
over a larger number of trials $n$ may help one achieve a target mean squared
error in location. An analytical derivation of the trade-off between $m$ and
$n$ in achieving a certain target MSE is relegated to future work. Further,
more sophisticated techniques for deriving lower bounds, such as Fano's
method~\cite[Sec.~2.7.1]{Tsybakov2009Introduction} are yet to be analyzed.

@@

While the bounds in Section~\ref{sec:lecam-lb} were derived for a
one-dimensional circular domain, extensions to a 1D linear domain (of fixed
length $S$) and to a $d$-dimensional domain (circular or linear) are fairly
straightforward, under certain conditions. For a 1D linear domain, we require
that sensors capture the entirety of the impulse response even at the edges of
the domain. Hence, sensors must be placed uniformly in $[-w/2, S{+}w/2]$. For
extending to $d$ dimensions, it suffices to apply Assouad's
Lemma~\cite[Sec.~2.7.2]{Tsybakov2009Introduction} to separate the problem into
$d$ independent 1D problems. Such extensions substantially broaden the
application domains where this bound could be useful.

@@

%So far, we've relied on the presence of a circular domain for to make the
%arguments in the proof. However, the most important requirement was that
%\emph{all} parts of the impulse response should get \emph{sensed} uniformly.
%So, if $\theta$ were restricted to the region $[0, S)$ (now on a linear,
%non-periodic domain), and if sensors were available to sense the signal from
%$[-w/2, S{+}w/2)$ (for an impulse response with restricted support as described
%in Section~\ref{sec:signal-model}), then the proof for sensor noise would still
%fall through, albeit with slightly different constants.
